Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-07 paper/paper.tex
first major revision

\documentclass[review,onefignum,onetabnum]{siamart190516}
%\usepackage{natbib}
%\usepackage[sort]{cite}
\pdfoutput=1
%\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[framemethod=TikZ]{mdframed}
\usetikzlibrary{shapes}
\usetikzlibrary{snakes}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithmic}
\usepackage[linesnumbered, ruled, vlined, algo2e]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

\title{Solving the independent set problem by generic programming tensor networks
\thanks{\funding{...}}
}

@@

\begin{abstract}
	This paper is about solving the independent set problem by eincoding this problem to a tensor network. 
    We show how to obtain the maximum independent set size, the independence polynomial and optimal configurations of a graph by engineering the tensor element algebra.
\end{abstract}

% REQUIRED
\begin{keywords}
  maximum independent set, tensor network
\end{keywords}

@@

\section{Introduction}
\blue{some literature does not distinguish maximal and maximum, efficient counting maximal independent sets~\cite{Manne2013}, overlap gap property~\cite{Gamarnik2013, Gamarnik2019}, independence polynomial at -1~\cite{Bousquet2008,Levit2009}, treewidth of 3-regular graphs ~\cite{Fomin2006}}
In this work, we introduce a tensor based framework to study the famous graph problem of finding independent sets.
Given an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set that for any $u,v \in I$, there is no edge connecting $u$ and $v$ in $G$.
The problem of finding the maximum independent set (MIS) size $\alpha(G) \equiv \max_{I}|I|$ belongs to the complexity class NP-complete~\cite{Hastad1996}, which is unlikely to be decided in polynomial time.
It is hard to even approximate this size in polynomial time within a factor $|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$.
The exhaustive search for a solution costs time $2^{|V|}$.
More efficient algorithms to compute the MIS size exactly includes the branching algorithm~\cite{Tarjan1977,Robson1986} and dynamic programming.
Without changing the fact of exponential scaling in computing time, the branching algorithm gives a smaller base.
For example, in ~\cite{Xiao2017}, a sophisticated branching algorithm has a time complexity $1.1996^n n ^{O(1)}$.
The dynamic programming approach~\cite{Courcelle1990,Fomin2013} works better for sparse graphs with small tree width $tw(G)$~\cite{Fomin2006}, it gives an algorithms of complexity $O(2^{tw(G)}tw(G)n)$.
People are interested in solving the independent set problem better not only because it is an NP-complete problem that directly related to other NP-complete problems like maximal cliques and vertex cover~\cite{Moore2011},
but also for its close relation with physical applications like hard spheres lattice gas model~\cite{Dyre2016}, and Rydberg hamiltonian~\cite{Pichler2018}.
However, in these applications, knowing the MIS size and one of the optimum solutions is not the only goal.
%A set is independent if and only if it is a clique in the graph's complement, so the two concepts are complementary.
%A set is independent if and only if its complement is a vertex cover. Therefore, the sum of the size of the largest independent set $\alpha (G)$ and the size of a minimum vertex cover $\beta (G)$ is equal to the number of vertices in the graph.
People often ask different questions about independent sets in order to understand the landscape of their models better.
These questions includes but not limited to, counting all independent sets, obtaining all independent sets of size $\alpha(G)$ and $\alpha(G)-1$,
counting independent sets of different sizes, and understanding the effect of a local gadget.
In this work, we attack this problem by mapping it to an generic tensor network.
It does not give a better time complexity comparing to dynamic programming, but is versatile enough to answer the above questions by engineering the tensor elements with minimum effort.

\section{Tensor networks}
%The word ``einsum'' is a shorthand for Einstein's summation, however, modern einsum notation in program is actually invented by a group of programmers.
A tensor network can be viewed as a generalization to of binary matrix multiplication to n-ary tensor contraction.
Let $A, B$ be two matrices, the matrix multiplication is defined as $C_{ik} = \sum_{j}A_{ij}B_{jk}$.
A traditional tensor network refers to the Einstein's notation. In this notation, the matrix multiplication is denoted as $C_i^k = A_i^j B_j^k$, where the paired subscript and superscript $j$ is a dummy index summed over, hence each index appears precisely twice.
When we have multiple tensors doing the above sum-product operation, we get a traditional tensor network~\cite{Orus2014}.
A traditional tensor network can be represented as a mutigraph with open edges by viewing a tensor as a vertex, a label pairing two tensors as an edge, and the remaining unpaired labels as open edges.

\begin{example}
   A traditional tensor network $C_i^l = A_{ij}^{km} B^l_{km} V^j$ has the following multigraph representation.

\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
    \def\dx{0};
    \def\r{0.25cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white, fill=black, dot=2*\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white, fill=black, dot=2*\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white, fill=black, dot=2*\r] at (\cx+\dx,\cy) (c) {V};
    %\draw [black,thick] (\ax+\dx,\ay) .. controls (0.5*\ax+0.5*\bx+\dx+0.2,0.5*\ay+0.5*\by-0.2) .. (\bx+\dx,\by);
    %\draw [black,thick] (\ax+\dx,\ay) .. controls (0.5*\ax+0.5*\bx+\dx-0.2,0.5*\ay+0.5*\by+0.2) .. (\bx+\dx,\by);
    \draw [black,thick] (a) .. controls (0.4, 0.6) .. (b);
    \draw [black,thick] (a) .. controls (0.6, 0.4) .. (b);
    \draw [black,thick] (a) -- (c);
    \draw [black,thick] (a) -- (\ax+\dx-0.8,\ay);
    \draw [black,thick] (b) -- (\bx+\dx+0.8,\by);
    \node[color=black] at (0.5*\ax+0.5*\bx-0.3+\dx,0.5*\ay+0.5\by+0.3) {k};
    \node[color=black] at (0.5*\ax+0.5*\bx+0.3+\dx,0.5*\ay+0.5\by-0.3) {m};
    \node[color=black] at (0.5*\ax+0.5*\cx+\dx+0.2,0.5*\ay+0.5\cy+0.2) {j};
    \node[color=black] at (\ax+\dx-0.5,\ay-0.2) {i};
    \node[color=black] at (\bx+\dx+0.5,\by-0.2) {l};
\end{tikzpicture}}

\end{example}

Here, we want to use a generalized tensor network notation by not restricting the number of times a label appears in the notation,
hence whether an index is a superscript or a subscript makes no sense now.
It is also called einsum, sum-product network or factor graph~\cite{Bishop2006} in some contexts.
The graphical representation of a tensor network in this paper is a hypergraph, where an edge (label) can be shared by an arbitrary number of vertices (tensors).

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network, it represents $C_{ijk} = \sum_{ml}A_{jkm} B_{mia} V_{jm}$.
Its hypergraph representation is as the following, where we use different color to annotate different hyperedges.

@@

In the main text, we stick to the our generalized tensor network notation rather than the traditional notation.
As a note to those who are more familiar with the traditional tensor network representation,
although one can easily translate a generalized tensor network to the equivalent traditional tensor network by adding $\delta$ tensors (a generalization of identity matrix to higher order).
It can sometime increase the contraction complexity of a graph. We have an example demonstrating this point in \App{app:tensorbad}.

\section{Independence polynomial}
One can encode the independence polynomial~\cite{Levit2005, Ferrin2014, Harvey2017} of $G$ to a tensor network.
Independence polynomial is an important graph polynomial that contains the counting information of an independent set problem. It is defined as
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$, and $\alpha(G)$ is the maximum independent set size.
The problem of computing independence polynomial belongs to the complexity class \#P-hard.
A traditional approach to compute independence polynomial rigorusly requires a computing time $O(1.442^n)$~\cite{Ferrin2014}\blue{I am not sure about this complexity, this is baed on the naive analysis of theorem 2.2 in ~\cite{Ferrin2014}}.
%with which computing the independence polynomial for a square lattice of size $30\times 30$ would be impossible.
There are some interests in approximating this polynomial efficiently~\cite{Harvey2018}, but here, we focus on the rigorous approaches.
We encode this polynomial to a tensor network by placing a rank one tensor of size $2$ parametrized by $x_i$ on a vertex $i$
\begin{equation}
    W(x_i)_{s_i} = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right)_{s_i},\label{eq:wtensor}
\end{equation}
and a rank two tensor of size $2 \times 2$ on an edge $(i,j)$
\begin{equation}
    B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right)_{s_is_j},\label{eq:btensor}
\end{equation}
where a tensor index $s_i$ is a boolean variable that having the meaning of being 1 if vertex $i$ is in the independent set, 0 otherwise.
It corresponds to a hyperedge in the hypergraph.
The contraction of such a tensor network gives
\begin{equation}
    P(G, \{x_1,\ldots,x_n\}) = \sum\limits_{s_1, s_2, \ldots, s_n = 0}^{1} \prod\limits_{i=1}^n W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1,\ldots,s_n\}$ and accumulates the product of tensor elements to the scalar output $P$.
We can see an edge tensor represents the restriction on an edge that if both vertices connected by it are included in the set, then such configuration has no contribution to the output.
When we set $x_i = x$, the contraction result corresponds to the independence polynomial.
One can see the connection from the fact that the product over vertex tensor elements gives a factor $x^k$, where $k=\sum_i s_i$ counts the set size, 
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set, $0$ otherwise.
One directly benefit of mapping the independent set problem to a tensor network is one can take the advantage of recently developed techniques in tensor network based quantum circuit simulations ~\cite{Gray2021,Pan2021},
where people evaluate a tensor network by pairwise contracting tensors in a heuristic order.
A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{tw(G)})$.
Here $tw(G)$ is the tree width of the line graph of a tensor network hypergraph, while the line graph of a tensor network hypergraph corresponds to the original graph $G$ that we mapped from.~\cite{Markov2008}
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up our computation for certain tensor element types.

\begin{example}
    Mapping a graph (left) to a tensor network, the resulting tensor network is shown in the right panel.
    In the generalize tensor network's graphical representation, a vertex is mapped to a hyperedge,
    and an edge is mapped to an edge tensor.

    \centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]

>>>>

\documentclass[onefignum, onetabnum]{siamart190516}
%\usepackage{natbib}
%\usepackage[sort]{cite}
\pdfoutput=1
%\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[framemethod=TikZ]{mdframed}
\usetikzlibrary{shapes}
\usetikzlibrary{snakes}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithmic}
\usepackage[linesnumbered, ruled, vlined, algo2e]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
%\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

\title{Computing properties of independent sets by generic programming tensor networks
\thanks{\funding{...}}
}

@@

\begin{abstract}
We introduce a new approach to compute various properties of independent sets by encoding the problems into tensor networks with different tensor element algebra. The type of graph problems that can be computed using this method includes: independence number, the number of independent sets, independence polynomial, maximal independence polynomial, and enumeration of the maximum independence set configurations. Although the computational complexity inevitably scales exponentially with the graph size, the performance of our method using tensor network contraction with generic programming is on par or outperforms state-of-the-art, sophisticated methods; on the other hand, our algorithms are very simple to implement and one can directly utilize recent advances in tensor network contraction techniques. To demonstrate the versatility of this tool, we apply it to a few examples including the calculations of the hard-square constant, Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs.
\end{abstract}

% REQUIRED
\begin{keywords}
independent set, tensor network, maximum independent set, independence polynomial
\end{keywords}

@@

\section{Introduction}
\blue{efficient counting maximal independent sets~\cite{Manne2013}}
In graph theory and combinatorics, there are many interesting but hard computational problems concerning properties of independent sets. An independent set is a set of vertices in a graph where no two vertices are adjacent to each other. More formally, given an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set that for any $u,v \in I$, there is no edge connecting $u$ and $v$ in $G$. The problem of finding an independent set of the largest possible size, the maximum independent set (MIS), is a paradigmatic NP-complete problem~\cite{Hastad1996}. The size of the MIS of a graph $G$ is called the independence number, denoted as $\alpha(G) \equiv \max_{I}|I|$.

Finding $\alpha(G)$ is a hard computational problem; moreover, it is NP-hard to even approximate $\alpha(G)$ within a factor $|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$. Naive exhaustive search for the MIS requires time $O(2^{|V|})$. More efficient exact algorithms include the branching algorithms and dynamic programming. Sophisticated branching algorithms can reduce the base of the exponential time scaling to, e.g., $1.1893^n n ^{O(1)}$~\cite{Xiao2017} \red{reading the abstract, I think the algorithm is $1.1996^{|V|} {|V|}^{O(1)}$. The $1.1893^{|V|} {|V|}^{O(1)}$ scaling is for graphs with maximum degree 6}. Dynamic programming approaches~\cite{Courcelle1990, Fomin2013} works better for graphs with a small treewidth $\text{tw}(G)$; these methods can produce algorithms of complexity $O(2^{\text{tw}(G)} \text{tw}(G) |V|)$.

The independent set problem is closely related to the clique problem and the vertex cover problem~\cite{Moore2011}; more concretely, an MIS of a graph $G$ corresponds to a maximum clique in the complement graph of $G$, and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$. Together, these problems find a wide range of applications in scheduling, logistics, social network analysis, bioinformatics, wireless networks and telecommunication, map labelling, computer vision, etc.~\cite{Butenko2003Maximum, Wu2015review}. Besides the standard MIS problem, there are many other interesting problems pertaining to independent sets, such as the number of independent sets, independence polynomial, maximal independence polynomial, and enumeration of optimal and sub-optimal configurations. Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007Monte} in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready}; they can, for example, be used to understand phase transitions and to identify harder graphs in an ensemble of graphs \red{cite experiment}.

In this work, we introduce a tensor-network based framework to compute the various properties pertaining to independent sets. We map different problems into generic tensor network contraction with various tensor element algebra. Our approach does not necessarily provide a better time complexity compared to dynamic programming, but the tensor network approach is highly versatile so different problems can be generically solved with minimal implementation efforts. We benchmark our algorithms, the implementation of which benefits from recent advances in tensor-network contraction for the purposes of quantum circuit simulations~\cite{Gray2021, Pan2021}. Lastly, we provide a few examples and show that the toolbox can be used to calculate the hard-square constant, Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs.

\section{Tensor networks}\red{I removed the superscript and subscript, since I think it's not necessary to introduce those.}
A tensor network~\cite{Cirac2020Matrix, Orus2014} can be viewed as a generalization of matrix multiplication to multi-dimensional tensor contraction. Let $A, B$ be two matrices; the matrix multiplication is defined as $C_{ik} = \sum_{j}A_{ij}B_{jk}$. Using the Einstein notation, we can write $C_{ik} = A_{ij}B_{jk}$, where $j$ is implicitly summed over. In the standard notation, each index can appear at most twice. A tensor network consists of multiple tensors performing the sum-product operation, i.e., contraction, and the tensor network can be represented as a multigraph with open edges by viewing a tensor as a vertex, a label pairing two tensors as an edge, and the remaining unpaired labels as open edges.

\begin{example}
   A tensor network $C_{il} = A_{ijkm} B_{kml} V_{j}$ has the following multigraph representation.

\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]
    \def\dx{0};
    \def\r{0.25cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white, fill=black, dot=2*\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white, fill=black, dot=2*\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white, fill=black, dot=2*\r] at (\cx+\dx,\cy) (c) {V};
    %\draw [black,thick] (\ax+\dx,\ay) .. controls (0.5*\ax+0.5*\bx+\dx+0.2,0.5*\ay+0.5*\by-0.2) .. (\bx+\dx,\by);
    %\draw [black,thick] (\ax+\dx,\ay) .. controls (0.5*\ax+0.5*\bx+\dx-0.2,0.5*\ay+0.5*\by+0.2) .. (\bx+\dx,\by);
    \draw [black,thick] (a) .. controls (0.4, 0.6) .. (b);
    \draw [black,thick] (a) .. controls (0.6, 0.4) .. (b);
    \draw [black,thick] (a) -- (c);
    \draw [black,thick] (a) -- (\ax+\dx-0.8,\ay);
    \draw [black,thick] (b) -- (\bx+\dx+0.8,\by);
    \node[color=black] at (0.5*\ax+0.5*\bx-0.3+\dx,0.5*\ay+0.5\by+0.3) {k};
    \node[color=black] at (0.5*\ax+0.5*\bx+0.3+\dx,0.5*\ay+0.5\by-0.3) {m};
    \node[color=black] at (0.5*\ax+0.5*\cx+\dx+0.2,0.5*\ay+0.5\cy+0.2) {j};
    \node[color=black] at (\ax+\dx-0.5,\ay-0.2) {i};
    \node[color=black] at (\bx+\dx+0.5,\by-0.2) {l};
\end{tikzpicture}}
\end{example}

One can generalize the tensor network notation by removing the restriction that each index can appear at most twice. The notation can be considered as a generalized Einstein notation, which is sometimes called einsum, sum-product network or factor graph~\cite{Bishop2006} in different contexts. The graphical representation of a tensor network we use in this paper is a hypergraph, where an edge (label) can be shared by an arbitrary number of vertices (tensors).

\begin{example}
$C_{ijkl} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijkl} = \sum_{m}A_{jkm} B_{mil} V_{jm}$  \red{I've changed this. please check. The original was $C_{ijk} = A_{jkm} B_{mil} V_{jm}$ and $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$}. Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges.

@@

In the main text, we use the generalized tensor network notation. The generalized tensor network can be easily translated to the standard notation by adding identity tensors denoted as $\delta$ tensors. The example above can be written as $C_{ijkl} = A_{jkm} B_{mil} V_{jm} = A_{ukm} B_{pil} V_{vq} \delta_{mpq} \delta_{juv}$ in the standard notion. However, by adding the $\delta$ tensors, one may unnecessarily increase the contraction complexity of a graph, which we illustrate in \App{app:tensorbad}.

\section{Generic programming}\label{sec:generic}
\red{I suggest we have a section on generic programming, putting both the semiring stuff and the summary table to this section, since generic programming is one of the highlights of this paper as well. I think it doesn't need to be long. Jinguo, could you write one or two paragraphs introducing generic programming?}

\red{Generic programming is ...}
\red{dispatch to different types etc. ...}

In the spirit of generic programming, the elements of the tensors in the tensor network do not always need to be floating point numbers or integers. In fact, for the tensor network contraction to make sense, one only needs the tensor elements to form a commutative semiring. A semiring is a ring without additive inverse, while a commutative semiring is a semiring where the multiplication operation is commutative. To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial, the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS configurations of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}. Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

\begin{table}[t!]\centering
\begin{minipage}{0.9\columnwidth}
\ra{1.3}
    \scalebox{1.0}{
        \begin{tabularx}{\textwidth}{bb}\toprule
            \hline
            \textbf{element type}     & \textbf{purpose} \\
            {real number}     & {counting all independent sets} \\
            {tropical number} (\Eq{eq:countingtropical})    & {finding the independence number} \\
            {tropical number with counting} (\Eq{eq:countingtropical})     & {finding the independence number and the number of MISs} \\
            {tropical number with configurations} (\Eq{eq:singleconfig})     & {finding the independence number and one MIS} \\
            {tropical number with sets} (\Eq{eq:set})    & {finding the independence number and enumeration of all MISs} \\
            {polynomial} (\Eq{eq:polynomial})     & {computing the independence polynomial exactly} \\
            {truncated polynomial} (\Eq{eq:max2poly})     & {counting MISs and independent sets of size $\alpha(G)-1$} \\
            {complex number}     & {fitting the independence polynomial with fast Fourier transform} \\
            {finite-field algebra (\Eq{eq:finitefield})} & {fitting the independence polynomial exactly using number theory} \\
            \bottomrule
        \end{tabularx}
    }
    \caption{Tensor element types and their purposes in calculating various independent set properties.}\label{tbl:generictypes}
\end{minipage}
\end{table}

\section{Independence polynomial and maximal independence polynomial}
\subsection{Independence polynomial}\label{sec:indpoly}
The independence polynomial is an important graph polynomial that contains the counting information of independent sets. It is defined as
\begin{equation}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$, and $\alpha(G)$ is the independence number. The total number of independent sets is thus equal to $I(G, 1)$. The problem of computing the independence polynomial belongs to the complexity class \#P-hard. One approach to exactly compute the independence polynomial requires a computational time of $O(1.442^{|V|})$~\cite{Ferrin2014}\blue{I am not sure about this complexity, this is based on the naive analysis of theorem 2.2 in ~\cite{Ferrin2014}}. There are some interesting works on efficiently approximating the independence polynomial~\cite{Harvey2018}, but in this work, we focus on exact methods to compute the independence polynomial.

The independence polynomial is closely related to the matching polynomial, the clique polynomial, and the vertex cover polynomial. In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{levit2005Independence}. Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{levit2005Independence}. We hope our algorithms to calculate the independence polynomial can also help further research in these directions.

To compute the independence polynomial of a graph $G$, we encode it to a tensor network. On the vertex $i$ of the graph $G$, we place a rank-one vertex tensor, $W(x_i)$, of size $2$ parametrized by $x_{i}$ and a rank-two edge tensor, $B$, of size $2 \times 2$ on an edge $(i,j)$
\red{I removed the indices $s_{i}$ etc, on the right of the equations, since that looks confusing to me. If I understand correctly, $s_{i}, s_{j}$ just denotes the index of the vector/matrix.}
\begin{equation}
    W(x_i)_{s_i} = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right),
    \qquad \quad 
       B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation}
where the tensor index $s_i$ is a boolean variable; $s_{i} = 1$ corresponds that the vertex $i$ is in the independent set, and $s_{i} = 0$ means otherwise. Let us assume the tensor elements are real numbers for now. The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$. The contraction of the tensor network representing $G$ gives
\begin{equation}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the scalar output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example). In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial. The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size, and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. In the implementation of the algorithm, we benefit from recent advances in tensor network contraction, where researchers working on quantum circuit simulation evaluate the tensor network by pairwise contracting tensors in a good heuristic order~\cite{Gray2021, Pan2021}. A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{\text{tw}(G)})$~\cite{Markov2008}. The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

\begin{example}\label{eg:tensorcontraction}
Mapping a graph (left) to a tensor network (right) that encodes the independence polynomial. In the generalized tensor network's graphical representation, a vertex is mapped to a hyperedge, and an edge is mapped to an edge tensor. \red{a vertex is mapped to a hyperedge?}

    \centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
        \def\dx{0};
        \def\r{0.25cm}
        \filldraw[fill=black] (\dx,0) circle [radius=\r];
        \filldraw[fill=black] (\dx,1.5) circle [radius=\r];
        \filldraw[fill=black] (1.5+\dx,0) circle [radius=\r];
        \filldraw[fill=black] (1.5+\dx,1.5) circle [radius=\r];
        \filldraw[fill=black] (2.5+\dx,2.5) circle [radius=\r];
        \draw [black,thick] (\dx,0) -- (\dx,1.5);
        \draw [black,thick] (\dx,0) -- (1.5+\dx,0);
        \draw [black,thick] (\dx,1.5) -- (1.5+\dx,1.5);
        \draw [black,thick] (1.5+\dx,0) -- (1.5+\dx,1.5);
        \draw [black,thick] (1.5+\dx,0) -- (\dx,1.5);
        \draw [black,thick] (2.5+\dx,2.5) -- (1.5+\dx,1.5);
        \node[color=white] at (\dx,0) {a};
        \node[color=white] at (\dx,1.5) {b};
        \node[color=white] at (1.5+\dx,0) {c};
        \node[color=white] at (1.5+\dx,1.5) {d};
        \node[color=white] at (2.5+\dx,2.5) {e};
        \def\dx{5};
        \def\r{0.25cm}
        \foreach \x/\y/\e in {0.75/0/ac, 0/0.75/ab, 1.5/0.75/cd, 0.75/1.5/bd, 0.75/0.75/bc, 2/2/de}
            \node[color=white,fill=black,dot=2*\r] at (\x+\dx,\y) (\e) {\scriptsize B};
        \foreach \x/\y/\v in {0/0/a, 0/1.5/b, 1.5/0/c, 1.5/1.5/d, 2.5/2.5/e}
            \node[color=black] at (\x+\dx,\y) (\v) {$s_\v$};
        \foreach \x/\y/\v in {-0.5/-0.5/a, -0.5/2.0/b, 2.0/-0.5/c, 2.0/1.0/d, 3.0/2.0/e}
            \node[color=white,fill=black,dot=\r] at (\x+\dx,\y) (\v\v) {};
        \foreach \x/\y/\v in {-0.5/-0.5/a, -0.5/2.0/b, 2.0/-0.5/c, 2.0/1.0/d, 3.0/2.0/e}
            \node[color=black] at (\x+\dx+0.5,\y) {\scriptsize $W(x_\v)$};
        \draw [cyan,thick] (a) -- (aa);
        \draw [cyan,thick] (a) -- (ab);
        \draw [cyan,thick] (a) -- (ac);
        \draw [blue,thick] (b) -- (bb);
        \draw [blue,thick] (b) -- (ab);
        \draw [blue,thick] (b) -- (bc);
        \draw [blue,thick] (b) -- (bd);
        \draw [red,thick] (c) -- (cc);
        \draw [red,thick] (c) -- (ac);
        \draw [red,thick] (c) -- (bc);
        \draw [red,thick] (c) -- (cd);
        \draw [green,thick] (d) -- (dd);
        \draw [green,thick] (d) -- (bd);
        \draw [green,thick] (d) -- (de);
        \draw [green,thick] (d) -- (cd);
        \draw [orange,thick] (e) -- (ee);
        \draw [orange,thick] (e) -- (de);
    \end{tikzpicture}}
